Geometric variational problems

### 2020-12-23 NSF2020/Description/preliminarywork.tex
left overs from NSF2020 proposal

\begin{wraptable}[6]{r}{0.55\textwidth}
\vspace{-0.8cm}
\centering{
  \begin{tabular}{|l|r|r|p{0.1\textwidth}|}
    \hline
          & HAP                     & experiment          & ref.\\
    \hline 
    $\KB$ & $11 \pm 2.5$ \kBT       & 10 \kBT               & \cite{Naetal15,VeBrPa15,NAGLE2000159,PhysRevLett.113.248102}\\
    \hline 
    $\KA$ & $34$ \kBT \; nm$^{-2}$  & 30--40 \kBT nm$^{-2}$ & \cite{Nagle17, Nagle17-2}\\
    \hline 
    $\KTH$ & $12$ \kBT \; nm$^{-2}$ & 10 \kBT nm$^{-2}$     & \cite{KUZMIN2005, KoNa15} \\ \hline
  \end{tabular}
}
  \vspace{-8pt}
\caption{\label{tab:moduli} \footnotesize Comparison of values of elastic
  moduli from the experimental literature and derived by HAP
  simulation.} 
\end{wraptable}
\section{Preliminary work}
\label{sec:preliminary_work}
PIs RR and YNY started work on HAP modeling with Szu-Pei Fu (then PI YNY's student and now PI RR's postdoc) and collaborators in 2017. 
%
%
%The HAP project was initiated in 2017 by PI RR and YNY. The lead author
%on our first paper~\cite{Fu2018_SIAM} was Szu-Pei Fu, who was PI YNY's
%PhD student and is currently a postdoc with PI RR. The collaboration
%involved theoretical work and simulation. The theoretical work included
%model development, proving the first variation formula~\eqref{stress}
%and force-free law, and an energy-based uniqueness theorem for the
%exterior domain problem~\eqref{SL}. 
%Coauthors Andreas Kl\"ockner
%(Computer Science, UIUC, NSF CAREER \#1654756) and his PhD student Matt
%Wala provided their Quadrature by Expansion, or QBX algorithm and Fu ran
%simulations on NJIT clusters.
%
%\begin{wraptable}[11]{l}{0.43\textwidth}
%\centering{
%  \begin{tabular}{|l|l|l|l|}
%    $n$   & (rel.~err.)$_F$
%    &  (rel.~err.)$_G$ \\
%    \hline
%% 32   &   4.07466     &   1.56437\\
%% 64   &   0.24664     &   0.03476\\
%% 128  &   0.00063     &   0.00006\\
%  32  & $4.07 \times 10^{+0}$ & $1.56 \times 10^{+0}$ \\
%  64  & $2.47 \times 10^{-1}$ & $3.48 \times 10^{-2}$ \\
%  128 & $6.30 \times 10^{-4}$ & $6.00 \times 10^{-5}$ 
%  \end{tabular}}
%\vspace{-5pt}
%\caption{\label{tab:spectral_force} 
%\footnotesize  Relative numerical errors (rel.~err.)$_F = \max_{i}
%  \|\mathbf{F}_i-\mathbf{F}_i^{\text{exact}}\|/\|\mathbf{F}_i^{\text{exact}}\|$
%  and (rel.~err.)$_G = \max_{i}
%  \|\mathbf{G}_i-\mathbf{G}_i^{\text{exact}}\|/\|\mathbf{G}_i^{\text{exact}}\|$
%  for force and torque respectively as a function of number of grid
%  points $n$ per particle.  The data is for $N = 5$ particles; two of
%  the particles are nearly touching at a distance 1\% of particle
%  diameter.} 
%\end{wraptable}
The results are summarized in \cite{Fu2018_SIAM}, where
small particle-number simulations successfully demonstrated gradient-driven self-assembly of
amphiphilic particles into two-dimensional micelles and bilayer membranes with realistic values of
the model parameters: decay length $\rho=2.5$ nm (based on measurements of
hydrophobic attraction between surfactant-coated
surfaces~\cite{Eriksson1989, Lin2005, Parsegian, Israelachvili80}),
particle diameter $\delta = 2.0$ nm (the physical monolayer thickness~\cite{Boal}),
and interfacial tension $\gamma = 4$ pN nm$^{-1}$ (based on
measurements for single-component bilayer lipid
membranes~\cite{GarciaSaez, KUZMIN2005, Petelska2012, Jackson2016}).

@@

\begin{wrapfigure}[22]{r}{0.5\textwidth}
\vspace{-5pt}
  \centering{\includegraphics[width=0.5\textwidth]{Figures/TankTreading.pdf}}
  \vspace{-25pt}
  \caption{\label{fig:tank_treading} \footnotesize
    Panel A: A two-dimensional, particle-based vesicle in a shear flow
    (arrows); $\mu = 1$ mPa s and $\dot{\gamma} = 0.5$ $\mu$s$^{-1}$
    Compare with~\cite{Brandner2019}).  Panel B: Angular velocities for
    the outer and the inner leaflets.  Panel C: The trajectory of the
    particle centers forms two ellipses with well-defined inclination
    angle $\Theta$. Panel D: The evolution of the inclination angle.
    Panels B and D share the same time axis.}
\end{wrapfigure}
We have simulated two-dimensional vesicles in shear flow
$\mathbf{u}_{\infty} = \dot{\gamma} x_2 \mathbf{i}_1$ with shear rate
$\dot{\gamma}$,
%(The symbol shear rate is unrelated to interfacial tension.)
and the particle-based system mimics the behavior observed in continuous
vesicles~\cite{torres-sanchez_millan_arroyo_2019,
mahapatra_saintillan_rangamani_2020, Steigmann99, C6SM02452A}. The basic
pattern is that of a vesicle bilayer approaching a steady, tank-treading
ellipse (Figure~\ref{fig:tank_treading}C). We analyzed the simulation
data in the context of theoretical work on
tank-treading~\cite{Finken2008, PhysRevLett.106.158103}.
Jeffrey's equation predicts a tank-treading frequency $-\dot{\gamma}/2$ (dashed line in
Figure~\ref{fig:tank_treading}B). In the particle simulation, we measured
a tank-treading frequency as the mean of the particle angular
velocities. The measured steady-state frequencies were within 7\% of the predicted 
frequency (dotted and solid curves in Figure~\ref{fig:tank_treading}B).
The inclincation angle of the particle-based simulation also
follows the inclination angle $\Theta(t)$ predicted by Jeffrey's
equation (Figure~\ref{fig:tank_treading}D), albeit with higher frequency oscillation
coming from particle reorganization.

@@

\begin{wrapfigure}[10]{l}{0.22\textwidth}
 \vspace{-3pt}
  \centering{\includegraphics[width=0.22\textwidth]{Figures/PW_fig5.pdf}}
 \vspace{-20pt}
  \caption{\label{fig:flow_rupture} \footnotesize Rupture of a
  two-dimensional vesicle at large shear rates.}
\end{wrapfigure}
The above considerations further justify using the particle-based model
to study material properties as outlined in
\S\ref{subsec:specific_aim_1}. Furthermore, the particle system allows
for intermonolayer slip~\cite{SHKULIPA2005823, ShkulipaThesis}. In
Figure \ref{fig:tank_treading}B, the inner angular velocity is slightly
less than the outer angular velocity. Intermonolayer slip enters
zero-thickness membrane models as a velocity jump boundary condition
\cite{schwalbe_vlahovska_miksis_2010}, for example, but in the present
model, it is a consequence of monolayer independence. For large shear
rates, viscous forces exceed the hydrophobic attraction leading to
 vesicle rupture~\cite{C8SM01501E}, at which point the flow carries away segmented membrane patches
(Figure \ref{fig:flow_rupture}).

>>>>

\begin{wraptable}[6]{r}{0.56\textwidth}
\vspace{-0.8cm}
\centering{
  \begin{tabular}{|l|r|r|p{0.1\textwidth}|}
    \hline
          & HAP                     & experiment          & ref.\\
    \hline 
    $\KB$ & $11 \pm 2.5$ \kBT       & 10 \kBT               & \cite{Naetal15,VeBrPa15,NAGLE2000159,PhysRevLett.113.248102}\\
    \hline 
    $\KA$ & $34$ \kBT \; nm$^{-2}$  & 30--40 \kBT nm$^{-2}$ & \cite{Nagle17, Nagle17-2}\\
    \hline 
    $\KTH$ & $12$ \kBT \; nm$^{-2}$ & 10 \kBT nm$^{-2}$     & \cite{KUZMIN2005, KoNa15} \\ \hline
  \end{tabular}
}
  \vspace{-8pt}
\caption{\label{tab:moduli} \footnotesize Comparison of values of elastic
  moduli from the experimental literature and derived by HAP
  simulation.} 
\end{wraptable}
\section{Preliminary work}
\label{sec:preliminary_work}
PIs RR and YNY started work on HAP modeling with Szu-Pei Fu (then PI YNY's student and now PI RR's postdoc) and collaborators in 2017. 
%
%
%The HAP project was initiated in 2017 by PI RR and YNY. The lead author
%on our first paper~\cite{Fu2018_SIAM} was Szu-Pei Fu, who was PI YNY's
%PhD student and is currently a postdoc with PI RR. The collaboration
%involved theoretical work and simulation. The theoretical work included
%model development, proving the first variation formula~\eqref{stress}
%and force-free law, and an energy-based uniqueness theorem for the
%exterior domain problem~\eqref{SL}. 
%Coauthors Andreas Kl\"ockner
%(Computer Science, UIUC, NSF CAREER \#1654756) and his PhD student Matt
%Wala provided their Quadrature by Expansion, or QBX algorithm and Fu ran
%simulations on NJIT clusters.
%
%\begin{wraptable}[11]{l}{0.43\textwidth}
%\centering{
%  \begin{tabular}{|l|l|l|l|}
%    $n$   & (rel.~err.)$_F$
%    &  (rel.~err.)$_G$ \\
%    \hline
%% 32   &   4.07466     &   1.56437\\
%% 64   &   0.24664     &   0.03476\\
%% 128  &   0.00063     &   0.00006\\
%  32  & $4.07 \times 10^{+0}$ & $1.56 \times 10^{+0}$ \\
%  64  & $2.47 \times 10^{-1}$ & $3.48 \times 10^{-2}$ \\
%  128 & $6.30 \times 10^{-4}$ & $6.00 \times 10^{-5}$ 
%  \end{tabular}}
%\vspace{-5pt}
%\caption{\label{tab:spectral_force} 
%\footnotesize  Relative numerical errors (rel.~err.)$_F = \max_{i}
%  \|\mathbf{F}_i-\mathbf{F}_i^{\text{exact}}\|/\|\mathbf{F}_i^{\text{exact}}\|$
%  and (rel.~err.)$_G = \max_{i}
%  \|\mathbf{G}_i-\mathbf{G}_i^{\text{exact}}\|/\|\mathbf{G}_i^{\text{exact}}\|$
%  for force and torque respectively as a function of number of grid
%  points $n$ per particle.  The data is for $N = 5$ particles; two of
%  the particles are nearly touching at a distance 1\% of particle
%  diameter.} 
%\end{wraptable}
The results are summarized in \cite{Fu2018_SIAM}, where
small particle-number simulations successfully demonstrated gradient-driven self-assembly of
amphiphilic particles into two-dimensional micelles and bilayer membranes with realistic values of
the model parameters: decay length $\rho=2.5$ nm (based on measurements of
hydrophobic attraction between surfactant-coated
surfaces~\cite{Eriksson1989, Lin2005, Parsegian, Israelachvili80}),
particle diameter $\delta = 2.0$ nm (the physical monolayer thickness~\cite{Boal}),
and interfacial tension $\gamma = 4$ pN nm$^{-1}$ (based on
measurements for single-component bilayer lipid
membranes~\cite{GarciaSaez, KUZMIN2005, Petelska2012, Jackson2016}).

@@

\begin{wrapfigure}[22]{l}{0.5\textwidth}
\vspace{-5pt}
  \centering{\includegraphics[width=0.49\textwidth]{Figures/TankTreading.pdf}}
  \vspace{-10pt}
  \caption{\label{fig:tank_treading} \footnotesize
    Panel A: A two-dimensional, particle-based vesicle in a shear flow
    (arrows); $\mu = 1$ mPa s and $\dot{\gamma} = 0.5$ $\mu$s$^{-1}$
    (compare with~\cite{Brandner2019}).  Panel B: Angular velocities for
    the outer and the inner leaflets.  Panel C: The trajectory of the
    particle centers forms two ellipses with well-defined inclination
    angle $\Theta$. Panel D: The evolution of the inclination angle.
    Panels B and D share the same time axis.}
\end{wrapfigure}
We have simulated two-dimensional vesicles in shear flow
$\mathbf{u}_{\infty} = \dot{\gamma} x_2 \mathbf{i}_1$ with shear rate
$\dot{\gamma}$,
%(The symbol shear rate is unrelated to interfacial tension.)
and the particle-based system mimics the behavior observed in continuous
vesicles~\cite{torres-sanchez_millan_arroyo_2019,
mahapatra_saintillan_rangamani_2020, Steigmann99, C6SM02452A}. The basic
pattern is that of a vesicle bilayer approaching a steady, tank-treading
ellipse (Figure~\ref{fig:tank_treading}C). We analyzed the simulation
data in the context of theoretical work on
tank-treading~\cite{Finken2008, PhysRevLett.106.158103}.
Jeffrey's equation predicts a tank-treading frequency $-\dot{\gamma}/2$ (dashed line in
Figure~\ref{fig:tank_treading}B). In the particle simulation, we measured
a tank-treading frequency as the mean of the particle angular
velocities. The measured steady-state frequencies were within 7\% of the predicted 
frequency (dotted and solid curves in Figure~\ref{fig:tank_treading}B).
The inclincation angle of the particle-based simulation also
follows the inclination angle $\Theta(t)$ predicted by Jeffrey's
equation (Figure~\ref{fig:tank_treading}D), albeit with higher frequency oscillation
coming from particle reorganization.

@@

\begin{wrapfigure}[10]{r}{0.22\textwidth}
 \vspace{-3pt}
  \centering{\includegraphics[width=0.21\textwidth]{Figures/PW_fig5.pdf}}
 \vspace{-7pt}
  \caption{\label{fig:flow_rupture} \footnotesize Rupture of a
  two-dimensional vesicle at large shear rates.}
\end{wrapfigure}
The above considerations further justify using the particle-based model
to study material properties as outlined in
\S\ref{subsec:specific_aim_1}. Furthermore, the particle system allows
for intermonolayer slip~\cite{SHKULIPA2005823, ShkulipaThesis}. In
Figure \ref{fig:tank_treading}B, the inner angular velocity is slightly
less than the outer angular velocity. Intermonolayer slip enters
zero-thickness membrane models as a velocity jump boundary condition
\cite{schwalbe_vlahovska_miksis_2010}, for example, but in the present
model, it is a consequence of monolayer independence. For large shear
rates, viscous forces exceed the hydrophobic attraction leading to
 vesicle rupture~\cite{C8SM01501E}, at which point the flow carries away
 segmented membrane patches (Figure \ref{fig:flow_rupture}).
